---
title: 'Rowvember'
author: 'Joseph V. Casillas'
date: '`r Sys.time()`'
output: 
  html_document: 
    highlight: textmate
    number_sections: yes
    theme: journal
    toc: yes
---

```{r, 'load-packages', echo=FALSE, message=FALSE, warning=FALSE}
library(plot3D)
library(plotly)
library(lme4)
library(lmerTest)
library(MuMIn)
library(DT)
library(stargazer)
library(here)
library(tidyverse)
library(magrittr)
library(knitr)
library(kableExtra)
library(magrittr)
```

```{r, 'import-clean-data', echo=FALSE, message=FALSE}
# Individual data
df_individual <- read_csv("./data/rowvember.csv")

# Add to dataframe
# df_individual <- rbind(df_individual, 
#   data_frame(rower = 'joseph', 
#              trial = 4, 
#              time = 2, 
#              distance = 663))

# Save to file
# write.table(df_individual, "./data/rowvember.csv", row.names = F, quote = F, sep = ",")

# Load all data
temp <- list.files(path = "./data/", full.names = TRUE, pattern = ".csv")
myfiles = lapply(temp[-1], read.csv, sep = ",")
df <- do.call("rbind", myfiles)

df <- mutate(df, 
             trial = as.factor(trial), 
             time = as.numeric(time), 
             distance = as.numeric(distance))
df <- df[-244, ]      
```

```{r, 'plot-setup', echo=F}
my_theme <- function() {
  theme_minimal() + 
    theme(
      plot.title = element_text(size = rel(1.5), face = "bold"), 
      plot.subtitle = element_text(size = rel(1)),
      plot.caption = element_text(color = "#777777", vjust = 0),
      axis.title = element_text(size = rel(1), hjust = 0.95, face = "italic"), 
      panel.grid.major = element_line(size = rel(.1), color = "grey70"), 
      panel.grid.minor = element_line(size = rel(.1), color = "grey70")
    )
}
```


# Analyses and plots

Each year in November we do a series of time trials in which we have to row 
for maximum distance in a set amount of time. The first week starts at 2 
minutes and the time is increased by 1 minute each week. For every time 
interval, each rower must complete 4 trials. This is the data from my 
rowvember's from 2016 through 2018.

## Total meters for Rowvember by year

Over the course of the month we complete `r 4*4` rows, so the meters really 
add up. Let's take a look at how much rowing actually goes down. 

```{r, 'total-meters-by-year-plot', echo=F, fig.width=5.5, fig.height=3, out.extra='style="float:left"'}
df_individual %>%
  group_by(rower, year) %>% 
  summarize(., total_meters = sum(distance)) %>%
  ggplot(., aes(x = year, y = total_meters)) + 
    geom_bar(stat = "identity", width = 0.05, fill = "grey60") + 
    geom_point(aes(fill = factor(year)), pch = 21, size = 8, 
               show.legend = F) + 
    scale_x_continuous(breaks = seq(2016, 2018, 1), 
                       labels = seq(2016, 2018, 1)) + 
	scale_fill_brewer(palette = "Set1") + 
    coord_cartesian(ylim = c(17100, 17700)) + 
    labs(y = "Total meters rowed", x = "Year", 
         title = "Rowvember analysis", 
         subtitle = "Total meters rowed by year") + 
    my_theme()
```

</br>
</br>
</br>

```{r, 'total-meters-by-year-table', echo=F}
# Total meters
df_individual %>%
  group_by(rower, year) %>% 
  summarize(., meters = sum(distance), 
               miles = (sum(distance) / 1609.344) %>% round(., 2)) %>%
  ungroup(.) %>% 
  select(-rower) %>% 
  add_row(., year = "total", 
             meters = sum(.$meters), 
             miles = sum(.$miles)) %>% 
  mutate(., year = if_else(year == "total", 
              cell_spec(year, background = "lightgrey", color = "black"), 
              cell_spec(year, background = "white")), 
            meters = if_else(meters > 20000, 
              cell_spec(meters, background = "lightgrey", color = "black"), 
              cell_spec(meters, background = "white")), 
            miles = if_else(miles > 20, 
              cell_spec(miles, background = "lightgrey", color = "black"), 
              cell_spec(miles, background = "white"))) %>% 
  kable(., format = "html", escape = F) %>% 
  kable_styling(., bootstrap_options = c("hover", "condensed", "responsive"), 
                   full_width = FALSE, position = 'center', font_size = 16) %>% 
  add_header_above(c(" ", "Totals" = 2))
```

</br>
</br>
</br>

In the last 3 years I have rowed a total of 
`r sum(df_individual$distance) %>% format(., scientific = F)` meters! That's 
`r (sum(df_individual$distance) / 1609.344) %>% round(., 2)` miles! 
One thing I was interested in was finding out if I was doing more work 
(rowing more distance) each year. Looking at the plot I see that I did indeed 
row more meters each year (the height of the colored dots increases). 
We can visualize the year to year improvement by subtracting the total meters 
of a given year from those of the previous year. 

```{r, 'total-improvement-by-year-plot', echo=F, fig.width=5.5, fig.height=3, out.extra='style="float:left"'}
df_individual %>%
  group_by(year) %>% 
  summarise(., total = sum(distance)) %>% 
  ungroup(.) %>% 
  spread(., year, total) %>% 
  mutate(., `16-17 difference` = `2017` - `2016`, 
            `17-18 difference` = `2018` - `2017`) %>% 
  select(., -`2016`, -`2017`, -`2018`) %>% 
  gather(., year, diff) %>% 
  ggplot(., aes(x = year, y = diff, fill = year)) + 
    geom_hline(yintercept = 0, lty = 3, color = "black") + 
    geom_bar(stat = "identity", width = 0.05, fill = "grey60", 
             position = position_dodge(0.5)) + 
    geom_point(pch = 21, size = 8, show.legend = FALSE) + 
    ylim(0, 300) + 
    scale_fill_brewer(palette = "Set1") + 
    labs(y = "Meters", 
         x = "", title = "Rowvember analysis", 
         subtitle = "Yearly improvement in total meters rowed") + 
    my_theme()
```

</br>
</br>
</br>
</br>

```{r, 'total-improvement-by-year-table', echo=F}
df_individual %>%
  group_by(year) %>% 
  summarise(., total = sum(distance)) %>% 
  ungroup(.) %>% 
  spread(., year, total) %>% 
  mutate(., `16-17` = `2017` - `2016`, 
            `17-18` = `2018` - `2017`) %>% 
  select(., -`2016`, -`2017`, -`2018`) %>% 
  gather(., Year, Improvement) %>% 
  kable(., format = "html") %>% 
  kable_styling(., bootstrap_options = c("hover", "condensed", "responsive"), 
                   full_width = FALSE, position = 'center', font_size = 16) 
```

</br>
</br>
</br>
</br>

This plot shows the difference in meters of total output from 2016 to 2017 and 
from 2017 to 2018. I've rowed more total meters every year. Specifically, I 
improved by 52 total meters from 2016 to 2017, and by 190(!) total meters from 
2017 to 2018. Nice. 

</br>
</br>

## Total meters per time interval

Rowvember can get pretty challenging because it involves 4 different row times. 
For me, the 2-minute and 3-minute rows are a sprint, which I prefer, but the 
3-minute row is torture. The 4 and 5-minute rows aren't easy, but I find them 
to be less difficult (and less stressful) because I just focus on holding my 
pace. Here I look at the total distance rowed each year as a function of the 
time inverval (2, 3, 4, or 5 minutes). 

```{r, 'total-meters-weekly-plot', echo=F, fig.width=5.5, fig.height=3, out.extra='style="float:left"'}
df_individual %>%
  group_by(., rower, year, time) %>%
  summarise(., total = sum(distance)) %>% 
  ggplot(., aes(x = year, y = total, dodge = factor(time))) + 
    geom_bar(stat = "identity", position = position_dodge(0.5), width = 0.2, 
             fill = "grey60") + 
    geom_point(aes(fill = factor(time)), position = position_dodge(0.5), 
               pch = 21, size = 7) + 
    coord_cartesian(ylim = c(0, 8000)) + 
    scale_fill_brewer(name = "Row time\n(minutes)", palette = "Set1") + 
    scale_x_continuous(breaks = seq(2016, 2018, 1), 
                       labels = seq(2016, 2018, 1)) + 
    labs(y = "Meters", 
         x = "", title = "Rowvember analysis", 
         subtitle = "Total meters rowed per time interval for 2016-2018") + 
    my_theme()
```

</br>

```{r, 'total-meters-weekly-table', echo=F}
# Total meters per week 
df_individual %>%
  group_by(., rower, year, time) %>%
  summarise(., total = sum(distance)) %>% 
  spread(., year, total) %>% 
  ungroup(.) %>% 
  select(-rower) %>% 
  rename(., "Row time" = time) %>% 
  kable(., format = "html") %>% 
  kable_styling(., bootstrap_options = c("hover", "condensed", "responsive"), 
                   full_width = FALSE, position = 'center', font_size = 16) %>% 
  add_header_above(c(" ", "Year" = 3))
```

</br>
</br>
</br>
</br>

It's pretty clear that with each additional minute of row time the total 
distance increases (big surprise). What may be more interesting is seeing 
the difference in total meters rowed per time interval for each year, so, 
again, we can subtract the total distance rowed from that of the previous 
year, for each time interval. 




```{r, 'total-improvement-by-interval-year-plot', echo=F, fig.width=6.25, fig.height=4, out.extra='style="float:left"'}

tibiyp_caption <- "2016/17 and 2017/18 comparison of total distances for 2, 3, 4, and 5 minute rows"

df_individual %>%
  group_by(year, time) %>% 
  summarise(., total = sum(distance)) %>% 
  ungroup(.) %>% 
  spread(., year, total) %>% 
  mutate(., `16-17 difference` = `2017` - `2016`, 
            `17-18 difference` = `2018` - `2017`) %>% 
  select(., -`2016`, -`2017`, -`2018`) %>% 
  gather(., year, diff, -time) %>% 
  ggplot(., aes(x = year, y = diff, fill = factor(time), dodge = factor(time))) + 
    geom_hline(yintercept = 0, lty = 3, color = "black") + 
    geom_bar(stat = "identity", width = 0.2, fill = "grey60", 
             position = position_dodge(0.5)) + 
    geom_point(pch = 21, size = 7, position = position_dodge(0.5)) +
    coord_cartesian(ylim = c(-50, 125)) + 
    scale_fill_brewer(name = "Row time\n(minutes)", palette = "Set1") + 
    labs(y = "Meters", 
         x = "", title = "Rowvember analysis", 
         subtitle = tibiyp_caption) + 
    my_theme()
```

</br>
</br>
</br>

```{r, 'total-improvement-by-interval-year-table', echo=F}
df_individual %>% 
  group_by(., rower, year, time) %>% 
  summarise(., total = sum(distance)) %>% 
  spread(., year, total) %>% 
  mutate(., `16-17` = `2017` - `2016`, 
            `17-18` = `2018` - `2017`) %>% 
  ungroup(.) %>% 
  select(., "Row time" = time, starts_with("1")) %>% 
  kable(., format = "html") %>% 
  kable_styling(., bootstrap_options = c("hover", "condensed", "responsive"), 
                   full_width = FALSE, position = 'center', font_size = 16) %>% 
  add_header_above(c(" ", "Comparison" = 2))
```

</br>
</br>
</br>
</br>
</br>

Text here












## Average meters per time interval

```{r}
df_individual %>% 
	group_by(., rower, time) %>% 
	summarize(., avg_distance = mean(distance), sd_distance = sd(distance)) %>% 
	mutate_if(., is.numeric, round, digits = 2) %>% 
	datatable(.)
```

## Fit model of meters as a function of time

```{r, echo=FALSE, output='hide'}
# Calculate mean value per time session and fit uncon. GCA 
# Create higher order orthogonal polynomial
df_individual <- as.data.frame(df_individual)
t <- poly(unique(df_individual$time), 2)
df_individual[df_individual$time == 2, 'ot1'] <- t[1, 1]
df_individual[df_individual$time == 3, 'ot1'] <- t[2, 1]
df_individual[df_individual$time == 4, 'ot1'] <- t[3, 1]
df_individual[df_individual$time == 5, 'ot1'] <- t[4, 1]
df_individual[df_individual$time == 2, 'ot2'] <- t[1, 2]
df_individual[df_individual$time == 3, 'ot2'] <- t[2, 2]
df_individual[df_individual$time == 4, 'ot2'] <- t[3, 2]
df_individual[df_individual$time == 5, 'ot2'] <- t[4, 2]

fit <- lmer(distance ~ (ot1+ot2) + year + 
           ((ot1+ot2) | trial) + 
           ((ot1 + ot2) | year), 
            data = df_individual, REML = FALSE)
summary(fit)
```

## Plots of meters as a function of time

```{r, fig.retina=2, fig.align='center', fig.width=8}
# Plot avg distance for each time 
df_individual %>%
  ggplot(., aes(x = time, y = distance, color = as.factor(year), dodge = year)) + 
  geom_point(position = position_dodge(0.5), size = 4, alpha = 0.3) + 
  stat_summary(fun.data = mean_sdl, geom = 'pointrange', 
  						 position = position_dodge(0.5)) + 
  stat_summary(fun.y = mean, geom = 'point', size = 2, 
  						 position = position_dodge(0.5)) + 
  stat_summary(aes(y = fitted(fit)), fun.y = mean, geom = 'line') + 
  scale_x_continuous(breaks = c(2, 3, 4, 5), 
                     labels = c('2min', '3min', '4min', '5min')) +
	ylim(500, 1700) + 
  xlab('Time interval') + ylab('Average distance (m)') + 
  theme_bw()
```




```{r, fig.height=7, fig.width=10, fig.retina=2, fig.align='center'}
# 3D plot 
with(df_individual, scatter3D(x = trial, y = time, z = distance, 
     colvar = distance, pch = 19, cex = 0.5, phi = 10, bty = 'g', 
     type = 'h', ticktype = "detailed", xlab = 'Trial', ylab = 'Time', 
     zlab = 'Distame (m)', zlim = c(0, 1800), xlim = c(0, 4.2), 
     ylim = c(1.8, 5.2), col = gg.col(100)))
```








</br></br></br></br></br></br></br></br></br>
</br></br></br></br></br></br></br></br></br>

# Group data

### Total meters (females)

```{r, echo=FALSE}
# Total meters (female)
df %>%
  filter(., group == 'female') %>%
  group_by(., athlete, group) %>%
  summarise(., totalM = sum(distance)) %>%
  arrange(., group, desc(totalM)) %>%
  as.data.frame(.) %>%
  datatable(.)
```

### Total meters (males)

```{r, echo=FALSE}
# Total meters (male)
df %>%
  filter(., group == 'male') %>%
  group_by(., athlete, group) %>%
  summarise(., totalM = sum(distance)) %>%
  arrange(., group, desc(totalM)) %>%
  as.data.frame(.) %>%
  datatable(.)
```

### Total meters as a function of time

```{r, echo=FALSE}
# Total meters by time
df %>%
  group_by(., athlete, group, time) %>%
  summarise(., totalM = sum(distance)) %>%
  arrange(., group, desc(totalM)) %>%
  by(., .[c("group", "time")], head, n = 10)
```

### Avg. meters by gender 

```{r, echo=FALSE}
# Avg meters by gender 
df %>%
  group_by(., athlete, group) %>%
  summarise(., totalM = sum(distance)) %>%
  group_by(., group) %>%
  summarise(., avg = round(mean(totalM), 2)) %>%
  datatable(.)
```

### Plot of total meters for top 10 overall

```{r, fig.align='center', fig.width=8, fig.retina=2, echo=FALSE}
# Plot total meters of top 10 overall
df %>%
  group_by(., athlete, group) %>%
  summarise(., totalM = sum(distance)) %>%
  arrange(., group, desc(totalM)) %>%
  by(., .["group"], head, n = 10) %>%
  Reduce(rbind, .) %>%
  ggplot(., aes(x = group, y = totalM, label = athlete, colour = group)) + 
  stat_summary(fun.data = mean_cl_boot, geom = 'errorbar', width = 0.2) + 
  stat_summary(fun.y = mean, geom = 'point', color = 'Black') + 
  geom_text(alpha = 0.4) + 
  scale_color_brewer(palette = 'Set1', guide = FALSE) + 
  theme_bw()
```

### Plot of top 10 in each trial 

```{r, fig.align='center', fig.width=10, fig.retina=2, echo=FALSE}
# Plot top 10 in each trial 
df %>%
  group_by(., athlete, group, trial, time) %>%
  arrange(., group, trial, time, desc(distance)) %>%
  by(., .[c("group", "trial", "time")], head, n = 10) %>%
  Reduce(rbind, .) %>%
  ggplot(., aes(x = trial, y = distance, shape = group, label = athlete)) + 
  facet_grid(. ~ time) + 
  stat_summary(aes(colour = group), fun.data = mean_cl_normal, geom = 'errorbar', width = 0.2) + 
  stat_summary(fun.y = mean, geom = 'point') + 
  geom_text(aes(colour = group), alpha = 0.3) + 
  scale_color_brewer(palette = 'Set1', guide = FALSE) + 
  theme_bw()
```

```{r, fig.align='center', fig.width=8, fig.retina=2, warning=FALSE, echo=FALSE, results='hide'}
# Build some models 
df$timeCentered <- df$time - mean(df$time)
df$trial <- as.numeric(df$trial)
fit_null <- lmer(distance ~ 1 + (time | athlete) + (trial | athlete), data = df, REML = FALSE)
fit_1    <- lmer(distance ~ 1 + time + (time | athlete) + (trial | athlete), data = df, REML = FALSE)
fit_2    <- lmer(distance ~ 1 + time + group + (time | athlete) + (trial | athlete), data = df, REML = FALSE)
fit_full <- lmer(distance ~ 1 + time * group + (time | athlete) + (trial | athlete), data = df, REML = FALSE)
anova(fit_null, fit_1, fit_2, fit_full, test = 'Chisq')
summary(fit_full)
r.squaredGLMM(fit_full)


df %>%
  ggplot(., aes(x = time, y = distance, colour = group, shape = group)) + 
  geom_jitter(alpha = 0.4) + 
  stat_summary(aes(y = fitted(fit_full)), fun.y = mean, geom = 'line') + 
  stat_summary(fun.data = mean_cl_boot, geom = 'pointrange') +
  stat_summary(fun.y = mean, geom = 'point', color = 'black') + 
  scale_color_brewer(palette = 'Set1') + 
  theme_bw()
```

```{r, fig.height=7, fig.width=10, fig.retina=2, fig.align='center', echo=FALSE}
with(df, scatter3D(x = trial, y = time, z = distance, 
     col.var = distance, pch = 19, cex = 0.5, phi = 10, bty = 'g', 
     type = 'p', ticktype = "detailed", xlab = 'Trial', ylab = 'Time', 
     zlab = 'Distame (m)', zlim = c(0, 1500), xlim = c(0, 4.2), 
     ylim = c(1.8, 5.2), col = gg.col(100)))
```




