---
title: 'Rowvember'
author: 'Joseph V. Casillas'
date: '`r Sys.time()`'
output: 
  html_document:
    number_sections: yes
    theme: cosmo
    toc: yes
---


# Load packages 



```{r, message=FALSE, warning=FALSE}
library(ggplot2); library(dplyr); library(tidyr) 
library(plot3D); library(plotly); library(lme4)
library(lmerTest); library(MuMIn); library(DT)
library(stargazer)
```


# Initial setup and data loading


```{r, eval=FALSE}
# clean working directory
rm(list = ls(all = TRUE))

# Set working directory
setwd("~/code/r/rowvember/")
```

```{r, echo=FALSE}
# Individual data
df_individual <- read.csv("./data/rowvember.csv", header = TRUE, quote = "")

# Add to dataframe
# df_individual <- rbind(df_individual, 
#   data.frame(rower = 'joseph', 
#              trial = 4, 
#              time = 5, 
#              distance = 1549))

# Save to file
# write.table(df_individual, "./data/rowvember.csv", row.names = F, quote = F, sep = ",")

# Load all data
temp <- list.files(path = "./data/", full.names = TRUE, pattern = ".csv")
myfiles = lapply(temp[-1], read.csv, sep = ",")
df <- do.call("rbind", myfiles)

df <- mutate(df, 
             trial = as.factor(trial), 
             time = as.numeric(time), 
             distance = as.numeric(distance))
df <- df[-244, ]      
str(df)
```



# Analysis and plots

## Individual data

### Total meters for November

```{r, results='asis'}
# Total meters
df_individual %>%
  summarize(., sum(distance)) %>%
  datatable(.)
```

### Total meters per week

```{r}
# Total meters per week 
df_individual %>%
  group_by(., time) %>%
  summarise(., sum(distance)) %>% 
  datatable(.)
```

### Fit model of meters as a function of time

```{r, echo=FALSE, output='hide'}
# Calculate mean value per time session and fit uncon. GCA 
# Create higher order orthogonal polynomial
df_individual <- as.data.frame(df_individual)
t <- poly(unique(df_individual$time), 2)
df_individual[df_individual$time == 2, 'ot1'] <- t[1, 1]
df_individual[df_individual$time == 3, 'ot1'] <- t[2, 1]
df_individual[df_individual$time == 4, 'ot1'] <- t[3, 1]
df_individual[df_individual$time == 5, 'ot1'] <- t[4, 1]
df_individual[df_individual$time == 2, 'ot2'] <- t[1, 2]
df_individual[df_individual$time == 3, 'ot2'] <- t[2, 2]
df_individual[df_individual$time == 4, 'ot2'] <- t[3, 2]
df_individual[df_individual$time == 5, 'ot2'] <- t[4, 2]

fit <- lmer(distance ~ (ot1+ot2) + ((ot1+ot2) | trial), data = df_individual, REML = FALSE)
summary(fit)
```

### Plots of meters as a function of time

```{r, fig.retina=2, fig.align='center', fig.width=8}
# Plot avg distance for each time 
df_individual %>%
  ggplot(., aes(x = time, y = distance)) + 
  stat_summary(fun.data = mean_se, geom = 'errorbar', width = 0.1) + 
  stat_summary(fun.y = mean, geom = 'point', color = 'darkred', size = 2) + 
  stat_summary(fun.y = mean, geom = 'point', color = 'darkgrey', size = 1.25) + 
  stat_summary(aes(y = fitted(fit)), fun.y = mean, geom = 'line') + 
  scale_x_continuous(breaks = c(2, 3, 4, 5), 
                     labels = c('2min', '3min', '4min', '5min')) +
  xlab('Time interval') + ylab('Average distance (m)') + 
  theme_bw()
```


```{r, fig.height=7, fig.width=10, fig.retina=2, fig.align='center'}
# 3D plot 
with(df_individual, scatter3D(x = trial, y = time, z = distance, 
     colvar = distance, pch = 19, cex = 0.5, phi = 10, bty = 'g', 
     type = 'h', ticktype = "detailed", xlab = 'Trial', ylab = 'Time', 
     zlab = 'Distame (m)', zlim = c(0, 1800), xlim = c(0, 4.2), 
     ylim = c(1.8, 5.2), col = gg.col(100)))
```









## Group data

### Total meters (females)

```{r, echo=FALSE}
# Total meters (female)
df %>%
  filter(., group == 'female') %>%
  group_by(., athlete, group) %>%
  summarise(., totalM = sum(distance)) %>%
  arrange(., group, desc(totalM)) %>%
  as.data.frame(.) %>%
  datatable(.)
```

### Total meters (males)

```{r, echo=FALSE}
# Total meters (male)
df %>%
  filter(., group == 'male') %>%
  group_by(., athlete, group) %>%
  summarise(., totalM = sum(distance)) %>%
  arrange(., group, desc(totalM)) %>%
  as.data.frame(.) %>%
  datatable(.)
```

### Total meters as a function of time

```{r, echo=FALSE}
# Total meters by time
df %>%
  group_by(., athlete, group, time) %>%
  summarise(., totalM = sum(distance)) %>%
  arrange(., group, desc(totalM)) %>%
  by(., .[c("group", "time")], head, n = 10)
```

### Avg. meters by gender 

```{r, echo=FALSE}
# Avg meters by gender 
df %>%
  group_by(., athlete, group) %>%
  summarise(., totalM = sum(distance)) %>%
  group_by(., group) %>%
  summarise(., avg = round(mean(totalM), 2)) %>%
  datatable(.)
```

### Plot of total meters for top 10 overall

```{r, fig.align='center', fig.width=8, fig.retina=2, echo=FALSE}
# Plot total meters of top 10 overall
df %>%
  group_by(., athlete, group) %>%
  summarise(., totalM = sum(distance)) %>%
  arrange(., group, desc(totalM)) %>%
  by(., .["group"], head, n = 10) %>%
  Reduce(rbind, .) %>%
  ggplot(., aes(x = group, y = totalM, label = athlete, colour = group)) + 
  stat_summary(fun.data = mean_cl_boot, geom = 'errorbar', width = 0.2) + 
  stat_summary(fun.y = mean, geom = 'point', color = 'Black') + 
  geom_text(alpha = 0.4) + 
  scale_color_brewer(palette = 'Set1', guide = FALSE) + 
  theme_bw()
```

### Plot of top 10 in each trial 

```{r, fig.align='center', fig.width=10, fig.retina=2, echo=FALSE}
# Plot top 10 in each trial 
df %>%
  group_by(., athlete, group, trial, time) %>%
  arrange(., group, trial, time, desc(distance)) %>%
  by(., .[c("group", "trial", "time")], head, n = 10) %>%
  Reduce(rbind, .) %>%
  ggplot(., aes(x = trial, y = distance, shape = group, label = athlete)) + 
  facet_grid(. ~ time) + 
  stat_summary(aes(colour = group), fun.data = mean_cl_normal, geom = 'errorbar', width = 0.2) + 
  stat_summary(fun.y = mean, geom = 'point') + 
  geom_text(aes(colour = group), alpha = 0.3) + 
  scale_color_brewer(palette = 'Set1', guide = FALSE) + 
  theme_bw()
```

```{r, fig.align='center', fig.width=8, fig.retina=2, warning=FALSE, echo=FALSE, results='hide'}
# Build some models 
df$timeCentered <- df$time - mean(df$time)
df$trial <- as.numeric(df$trial)
fit_null <- lmer(distance ~ 1 + (time | athlete) + (trial | athlete), data = df, REML = FALSE)
fit_1    <- lmer(distance ~ 1 + time + (time | athlete) + (trial | athlete), data = df, REML = FALSE)
fit_2    <- lmer(distance ~ 1 + time + group + (time | athlete) + (trial | athlete), data = df, REML = FALSE)
fit_full <- lmer(distance ~ 1 + time * group + (time | athlete) + (trial | athlete), data = df, REML = FALSE)
anova(fit_null, fit_1, fit_2, fit_full, test = 'Chisq')
summary(fit_full)
r.squaredGLMM(fit_full)


df %>%
  ggplot(., aes(x = time, y = distance, colour = group, shape = group)) + 
  geom_jitter(alpha = 0.4) + 
  stat_summary(aes(y = fitted(fit_full)), fun.y = mean, geom = 'line') + 
  stat_summary(fun.data = mean_cl_boot, geom = 'pointrange') +
  stat_summary(fun.y = mean, geom = 'point', color = 'black') + 
  scale_color_brewer(palette = 'Set1') + 
  theme_bw()
```

```{r, fig.height=7, fig.width=10, fig.retina=2, fig.align='center', echo=FALSE}
with(df, scatter3D(x = trial, y = time, z = distance, 
     col.var = distance, pch = 19, cex = 0.5, phi = 10, bty = 'g', 
     type = 'p', ticktype = "detailed", xlab = 'Trial', ylab = 'Time', 
     zlab = 'Distame (m)', zlim = c(0, 1500), xlim = c(0, 4.2), 
     ylim = c(1.8, 5.2), col = gg.col(100)))
```




